---
title: 'CohoBroodstock'
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  github_document:
    toc: yes
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "readme-figs/"
)
```

The goal of CohoBroodstock is to put a bunch of useful functions into one place to expedite
Libby's coho broodstock management work.

# Installing

If you don't already have the `devtools` package, then do:
```{r, eval=FALSE}
install.packages("devtools")
```

Then use it to install this from GitHub:
```{r, eval=FALSE}
devtools::install_github("eriqande/CohoBroodstock")
```

You also have to get the `related` package installed from R-forge.  That 
requires some gfortran compilation.  We did that for Libby already.

# Preparing Spawning matrices

This is set up now to use the _related_ package to compute the Rxy's.  
Input is a typical two-column format:  

- First column holds the individual IDs
- Every two columns after that are one locus.
- The file must have a row of column headers

The steps are:

1. first, make sure to load the package (and you might
as well load the tidyverse too...)
    ```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(CohoBroodstock)
    ```
    
1. read in the file and compute Rxy with computeRxy()
    ```{r computeRxy, cache=TRUE}
# get path to the example genotype file
# (typically you would pass it the path to your own file)
geno_file <- system.file("extdata",
                     "WSH_W1718_v5_two_column_data.txt.gz",
                     package = "CohoBroodstock")

# compute rxy.  This returns a tibble
rxy <- computeRxy(geno_file)
    ```
2. prepare the spawning matrix from the ouput of the last command using
spawning_matrix().  Like this:
    ```{r}
spawning_matrix(Rxy_tidy = rxy)
    ```
    
This creates, by default, two files named `spawn_matrix.csv` and `spawn_matrix_full.csv` in
the current working directory.



# Actual vs Optimal vs Random Relatedness

Here is how it goes. We will show it from the genotype stage.
```{r, message=FALSE, warning=FALSE}
geno_path <- system.file("extdata/IGH_W1819_geno_aor.txt", package = "CohoBroodstock")

rxys <- computeRxy(geno_path)
```

Here is what the first few rows of that look like
```{r}
rxys[1:10, ]
```


Then we need to read in the actual spawn pairs.  This should be two columns: the first one named Female and then Male:
```{r, message=FALSE}
pairs_file <- system.file("extdata/IGH--W1819--actual_spawn_pairs.csv", package = "CohoBroodstock")
actual_pairs <- read_csv(pairs_file)
```

This looks like this:
```{r}
actual_pairs[1:10, ]
```

Now, in order to use the function `aor_pairs()` we need to format the `rxys` tibble a little bit.
We need to only keep the individuals starting with "F" and those starting with "M", we need to
keep only comparisons between males and females, and we need to
name the columns "Female", "Male", and "rxy".  We do that with the `clean_computeRxy_output()` function.
```{r}
rxy_clean <- clean_computeRxy_output(rxys)
```
The result looks like this:
```{r}
rxy_clean[1:10,]
```

Before we feed these values int `aor_pairs` we have to remove Female F_7FN in `actual_pairs` because
it is named incorrectly (I think:  there is an F_07FN in the rxys file.  `aor_pairs` barks an error 
about that.)
```{r}
actual_pairs_corrected <- actual_pairs %>%
  filter(Female != "F_7FN")
```

Then we feed the actual values and all the values into the `aor_pairs` function.
```{r}
set.seed(10)  # set a random number seed for reproducibility
AOR <- aor_pairs(actual_pairs_corrected, rxy_clean)

# have a look at it:
AOR
```

Then plot those values in a histogram:
```{r aor_histo1}
cols <- c(Actual = "gold", Optimal = "limegreen", Random = "steelblue1")
ggplot(AOR, aes(x =  rxy, fill = `Spawn Pairs`)) +
  geom_histogram(position = "dodge", alpha = 0.75, binwidth = 0.03, color = "black", size = 0.2) +
  scale_fill_manual(values = cols)
```
