---
output:
  github_document:
    toc: yes
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# CohoBroodstock

The goal of CohoBroodstock is to put a bunch of useful functions into one place to expedite
Libby's coho broodstock management work.

## Realized vs Optimal vs Random Relatedness

Here is how it goes. First, read the symmetrical full matrix of rxy values:
```{r}
library(tidyverse)
file_path <- system.file("extdata/IGH_W1718_master_Kinsh_res.txt.gz", package = "CohoBroodstock")
rxys <- read_kinship_matrix(file_path)
```

Here is what the first few rows of that looks like
```{r}
rxys[1:10, ]
```


Then we need to read in the actual spawn pairs.  This should be two columns: first Female and then Male:
```{r}
pairs_file <- system.file("extdata/IGH_W1718_actual_spawn_pairs.csv", package = "CohoBroodstock")
actual_pairs <- read_csv(pairs_file)
```

This looks like this:
```{r}
actual_pairs[1:10, ]
```
